Demography and linked selection in wild and domesticated maize

### 2015-08-28 manuscript.tex
edits

@@

Our findings demonstrate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, and our analysis restricting to sites undergoing the strongest and weakest purifying selection confirm this result is robust to previously highlighted limitations \cite{enard2014}. Our approach does not address sweeps on noncoding variants, but it agrees with a growing body of previous evidence arguing against hard sweeps shaping maize variability. For instance, population genetic studies of local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} demonstrate the importance of selection on standing variation as opposed to on new mutations. Even during domestication, very few differences between maize and teosinte were fixed \cite{hufford2012}. Moreover, of domestication loci that have been well studied from a population genetics standpoint, only one shows a signal of a hard sweep \cite{wang2015}, although there are several consistent with selection on standing variation \cite{studer2011,wills2013} or selection on multiple mutations \cite{wills2013}.

>>>>

%JRI suggestions
%no strong evidence for hard sweeps, even considering constraint (GERP).
%although this is not super powerful test as misses sweeps on noncoding and only has power if substantial proportion of missense sites are swept, much previous evidence argues against hard sweeps in maize.  
%popgen studies of local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} show importance of standing variation. 
%even in domestication, few fixed differences between maize and teo \cite{hufford2012}, and of domestication loci that have been well studied from a popgen standpoint only one shows hard sweep \cite{wang2015} while there are multiple examples of standing variation \cite{studer2011, wills2013} or selection on multiple mutations \cite{wills2013}. \jri{maybe others? barren stalk 1 (gallavotti 2004)? ramosa 1 (sigmon 2010)? }
%
%this lack of hard sweeps patterning diversity differs from drosophila \citex and capsella \citex and many species with high Ne \jri{cite corbet-detig}. \jri{add something about corbet-dettig results for maize}
%but is consistent with humans \cite{hernandez2011}
%discussion of whether we expect hard vs soft in maize \cite{messer2013population}.
%$\theta_b = 2N\mu_b$ with $mu_b$ being mutation rate of beneficial alleles of selection coefficient s. 
%if $\theta_b > \frac{1}{log{Ns_b}}$ then adaptation from multiple mutations should be common. 
%if  $\theta_b > \frac{1}{log{2Ns_b}}$ then adaptation from standing neutral variation should be common, or if $\theta > \frac{1}{log(1+\frac{2Ns_b}{2Ns_d + 1}}$ for standing deleterious variation with selection coefficient $s_d$. \jri{i think this latter assumes codominance} codominant?)

%Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
%One known example of positive selection on a non-genic mutation involves the \emph{tb1} locus of maize, which is one of the best characterized examples of positive selection on a ``domestication gene'' in any crop \cite{clark2006}. However, the maize \emph{tb1} allele was already present in teosinte before domestication \cite{studer2011}, so this this example is in agreement with our finding that hard sweeps are rare. The \emph{gt1} locus is another well-characterized case of positive selection operating on standing variation at an enhancer region \cite{wills2013}. Instances of selection on standing variation such as these, often called soft sweeps, may be a major contributor to maize patterns of diversity \cite{beissinger2014}, and this could be part of the explanation as to why hard sweeps appear to be so rare, despite obvious morphological differences between maize and teosinte.

%Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}. However, we implemented a scan based on the H12 statistic, which is designed to identify both hard and soft sweeps with reasonable power \cite{garud2015}. Although the goal of this study was not to identify specific sweeps, it may be that the outlier sites distinguished by H12 are primarily composed of soft sweeps instead of hard. Similarly, previous studies scanning for evidence of positive selection in maize \cite{hufford2012} may have picked up primarily evidence of soft sweeps.

%We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution. In fact, the maize locus \emph{tga1} that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}. Surrounding \emph{tga1},  our data demonstrate a pattern consistent with a hard sweep. But, our data demonstrate that instances such as this are infrequent. This contrasts sharply with \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, where differences in diversity surrounding synonymous and non-synonymous substitutions are clear suggesting hard-sweeps are abundant. However, it agrees with what has been seen for humans \cite{hernandez2011}.

@@

Our findings demonstrate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show robust to concerns about power due to the effects of background selection  \cite{enard2014}. 
Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability. 
Population genomic studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation. 
Even among well-characterized domestication loci, only a single locus shows evidence of a hard sweep on a missense mutation \cite{wang2015}, but several are consistent with selection on standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
